\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-12 paper.qmd
no rigid page breaks allowed
Added description to the effect of density section

To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $7\text{-}D$ data having C-shaped structure is used. @fig-one-dens_clust-error a1 and b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used.

The models are fitted to these layouts. @fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space. In these projection the difference between the distribution of high-dimensional model error can be seen. The uniform densed C-shaped cluster

```{r}
#| label: fig-one-dens_clust-error
#| fig-height: 10
#| fig-width: 10
#| fig-pos: H
#| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error. The tSNE layouts of the C-shaped cluster, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed. Darker colors represent a high error, while lighter colors indicate a low error. The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end."

>>>>

To illustrate how tSNE organize different number of points within the structure in the tSNE layout, simulated $7\text{-}D$ data having C-shaped structure is used. @fig-one-dens_clust-error a1 and b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used. The \gD{} layouts are colored according to the high-dimensional model error. In b1, the large model error points are clustered in the top left corner, forming a separate cluster with high model errors.

The models are fitted to these layouts. @fig-one-dens_clust-error b1 and b2 show the fitted models in projections of the $7\text{-}D$ space. In these projections, the differences in the distribution of high-dimensional model errors can be observed. The uniform C-shaped cluster displays randomly distributed model errors, while the dense C-shaped cluster shows high model errors at the sparse end. In the dense C-shaped cluster, the high model errors occur because there are fewer data points positioned within the bins at the sparse end of the structure.

```{r}
#| label: fig-one-dens_clust-error
#| fig-height: 10
#| fig-width: 10
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "The sparse end of the structure exhibits a high model error, while the dense end shows a low model error. The tSNE layouts of the C-shaped structure, with uniform points (a1) and dense points (b1) are colored according to their high-dimensional error, along with their selected projections (a2 and b2), are displayed. Darker colors represent a high error, while lighter colors indicate a low error. The uniformly distributed points demonstrate random model errors throughout the structure, whereas the dense structure shows large model error at the sparse end. Videos of the langevitour animations are available at."

### 2025-01-12 paper.tex
no rigid page breaks allowed
Added description to the effect of density section

To illustrate how tSNE organize different number of points within the
structure in the tSNE layout, simulated \(7\text{-}D\) data having
C-shaped structure is used. Figure~\ref{fig-one-dens_clust-error} a1 and
b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used.

The models are fitted to these layouts.
Figure~\ref{fig-one-dens_clust-error} b1 and b2 show the fitted models
in projections of the \(7\text{-}D\) space. In these projection the
difference between the distribution of high-dimensional model error can
be seen. The uniform densed C-shaped cluster

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}

@@

\caption{\label{fig-one-dens_clust-error}The sparse end of the structure
exhibits a high model error, while the dense end shows a low model
error. The tSNE layouts of the C-shaped cluster, with uniform points
(a1) and dense points (b1) are colored according to their
high-dimensional error, along with their selected projections (a2 and
b2), are displayed. Darker colors represent a high error, while lighter
colors indicate a low error. The uniformly distributed points
demonstrate random model errors throughout the structure, whereas the
dense structure shows large model error at the sparse end.}

>>>>

To illustrate how tSNE organize different number of points within the
structure in the tSNE layout, simulated \(7\text{-}D\) data having
C-shaped structure is used. Figure~\ref{fig-one-dens_clust-error} a1 and
b1 shows \gD{} layouts for tSNE. The default hyper-parameters are used.
The \gD{} layouts are colored according to the high-dimensional model
error. In b1, the large model error points are clustered in the top left
corner, forming a separate cluster with high model errors.

The models are fitted to these layouts.
Figure~\ref{fig-one-dens_clust-error} b1 and b2 show the fitted models
in projections of the \(7\text{-}D\) space. In these projections, the
differences in the distribution of high-dimensional model errors can be
observed. The uniform C-shaped cluster displays randomly distributed
model errors, while the dense C-shaped cluster shows high model errors
at the sparse end. In the dense C-shaped cluster, the high model errors
occur because there are fewer data points positioned within the bins at
the sparse end of the structure.

@@

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}

@@

\caption{\label{fig-one-dens_clust-error}The sparse end of the structure
exhibits a high model error, while the dense end shows a low model
error. The tSNE layouts of the C-shaped structure, with uniform points
(a1) and dense points (b1) are colored according to their
high-dimensional error, along with their selected projections (a2 and
b2), are displayed. Darker colors represent a high error, while lighter
colors indicate a low error. The uniformly distributed points
demonstrate random model errors throughout the structure, whereas the
dense structure shows large model error at the sparse end. Videos of the
langevitour animations are available at.}
